Implicit spatial model to krill Dynamic Population in Antarctic Peninsula, 48-1 SubArea. SS3 aplications

### 2024-05-28 SS3_Krill_481.Rmd
improve Intro

```{r setup1, echo=FALSE}
rm(list = ls())
knitr::opts_chunk$set(eval=FALSE,
                      echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = 'center',
                      fig.pos = "H",
                      dev = 'jpeg',
                      dpi = 300,
                      tidy.opts=list(width.cutoff=50), 
                      tidy=TRUE)
#XQuartz is a mess, put this in your onload to default to cairo instead
options(bitmapType = "cairo") 
# (https://github.com/tidyverse/ggplot2/issues/2655)
# Lo mapas se hacen mas rapido
```

@@

The stock assessment model for *Euphausia superba* (hereafter krill) in the Antarctic Peninsula, operating within a spatial-temporal and ecosystem framework, highlights the importance of incorporating spatial heterogeneity. The spatial heterogeneity, such as variations in oceanographic conditions and habitat suitability across different regions, is likely to outperform a population dynamics of krill. By accounting for spatial heterogeneity, a integrated stock assessment model can better capture localized dynamics and interactions, leading to more accurate predictions of krill population dynamics. This nuanced approach allows for a more refined understanding of different zones shape krill abundance, fishing mortality and recruitment patterns across the Antarctic Peninsula. Consequently, by acknowledging and integrating spatial heterogeneity, the stock assessment model can provide more robust insights into the ecological dynamics of krill populations and improve CCAMLR fishery management and conservation strategies in this region.

@@

**(add refs)**

Antarctica, with its pristine and unique ecosystems, hosts one of the most crucial marine resources, the Antarctic krill (Euphausia superba). These small crustaceans play a pivotal role in the Southern Ocean's food web, serving as a primary food source for a wide range of marine species, including whales, seals, and seabirds. Given its ecological significance and economic value, effective management of krill stocks is paramount to ensure the sustainability of Antarctic ecosystems and support fisheries-dependent communities.

In this context, the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) plays a central role in the conservation and sustainable management of marine resources in the Southern Ocean. As part of its mandate, CCAMLR regularly conducts stock assessments to evaluate the status of key species, including Antarctic krill, and inform decision-making processes regarding fisheries management and conservation measures.

Traditionally, stock assessments have relied on single-source data, primarily from fisheries catch data or population monitoring surveys. However, these approaches may overlook critical aspects of krill ecology, such as spatial heterogeneity in abundance and distribution, as well as the influence of environmental factors and predator-prey interactions. Recognizing these limitations, there is a growing consensus on the need for integrated stock assessment models that incorporate multiple sources of information and consider the complex interplay of ecological variables.

In response to this need, we propose a comprehensive stock assessment of Antarctic krill in the Antarctic Peninsula, leveraging an integrated modeling approach. Unlike previous assessments, our model will integrate data from both fisheries and population monitoring efforts, providing a more holistic understanding of krill dynamics. Moreover, our approach will go beyond conventional methods by explicitly considering spatial heterogeneity in krill abundance and distribution across the Antarctic Peninsula.

Spatial heterogeneity is a fundamental aspect of krill ecology, driven by a myriad of factors including oceanographic conditions, sea ice dynamics, and predator foraging behavior. By incorporating spatially implicit information into our model, we aim to capture the complex spatial patterns of krill distribution and better understand the underlying drivers of variability in krill abundance. This spatially implicit approach will allow us to identify krill hotspots, areas of high productivity, and potential ecological corridors, providing valuable insights for conservation planning and management strategies.

@@

## Study Area in the Antarctic Peninsula

The study area for the assessment of krill populations in the Antarctic Peninsula is divided into strata to account for spatial heterogeneity. The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill. Therefore, dividing the study area into strata allows for a more detailed and accurate assessment of krill populations by considering local variations in environmental factors.

@@

In addition to existing data sources, researchers may submit requests to the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) Secretariat for access to additional data relevant to krill population assessment. These requests may include data from scientific research programs, environmental monitoring initiatives, and other sources that can contribute to improving the accuracy and reliability of the assessment.

(Figure \@ref(fig:path))

@@

Where:
- \(N_t\) is the abundance of krill at time \(t\).
- \(N_{t-1}\) is the abundance of krill at the previous time step.
- \(r\) is the intrinsic growth rate of the population.
- \(M\) is the mortality rate.
- \(R\) is the recruitment of new individuals into the population.

@@

The stock assesment model was configured using Stock Synthesis (SS3 hereafter)[SS3](https://vlab.noaa.gov/web/stock-synthesis) [@Methot2013; @methot2020stock] with the most updated version (V.3.30.21). SS3 is a structured age and size stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"*. SS3 has a stock population sub-model that simulates growth, maturity, fecundity, recruitment, movement, and mortality processes, and observation sub-models and expected values for different types of data. The model is coded in `C++` with estimation parameters enabled by automatic differentiation (ADMB) [@Fournier2012; @Methot2013]. The analysis of results and outputs uses R tools and the graphical interface of the *r4ss* library [@Taylor2019; @Winker2023].

@@

@@

In Table 1 we have ten scenarios to test different option in modeling about main consideration in assessment of krill population.

| Scenario | Description                                      |
|:---------:|:-------------------------------------------------|
|    s01     | Fishery, Predator, Survey, Environmental data agreggate (Whole 48.1) |
|    s1     | Fishery data (Length, Index, Catch) by strata  and Predator and Env data      |
|    s2    | Fishery and Survey (AMLR) data Length, Index, Catch by strata and Predator and Env data |
|    s3     | Same "s2" Without S-R relation         |
|    s4     | Same "s2" Ricker S-R relation estimated          |
|    s5     | Same "s2" BH S-R relation weak (0.9 steepness)     |
|    s6     | Same "s2" BH S-R relation strong (0.6 steepness)     |
|    s7     | Same "s2" BH S-R relation mid-strong (0.65 steepness) estimated     |

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, eval=TRUE}
# install.packages("devtools")
# devtools::install_github("r4ss/r4ss", ref="development")

# 
# install.packages("caTools")
# library("caTools")
# 
# install.packages("r4ss")
library(r4ss)
library(here)
#remotes::install_github("PIFSCstockassessments/ss3diags")
library(ss3diags)
library(kableExtra)
library(doParallel) # facilita la ejecución paralela en R
detectCores()
registerDoParallel(8)
library(ggpubr)
```

```{r fig.height=7}
SSplotData(base.model01, subplot = 1, 
           fleetnames = c("Fishery", "Survey", "Predator"),
           fleetcol = c("blue","green", "red"))

SSplotData(base.model2, 
           subplot = 1,
           pheight = 15)

```

Respecto a los valores y parametros biologicos modelados, los siguientes graficos identifican los estimadores puntuales del recurso

```{r}
SSplotSelex(base.model2,
            subplots = 1)
```

```{r eval=FALSE}
SSplotBiology(base.model2, 
              subplots =2, 
              labels = c("Length (cm)",
                         "Age (yr)",
                         "Maturity", 
                         "Mean weight (kg) in last year",
                          "Spawning output",
                         "Length (cm, beginning of the year)", 
                         "Natural mortality",
                          "Female weight (kg)", 
                         "Female length (cm)", 
                         "Fecundity", 
                         "Default fecundity label",
                          "Year", 
                         "Hermaphroditism transition rate", 
                         "Fraction females by age at equilibrium"),
 )

```

aporte de las cohortes por año para las capturas.
```{r eval=FALSE}
SSplotCohortCatch(base.model2, subplots = 1)
```

\quad

AJuste de tallas por flota
```{r eval=FALSE}
SSplotComps(base.model2, subplots = 1)
```

Otros plots
```{r}
SSplotDynamicB0(base.model2, uncertainty = F)
#SSplotSPR(base.model3)
```

```{r eval=FALSE}
SSplotPars(base.model2)

```

Salida de las biomasas con las dos flotas

```{r}
SSplotTimeseries(base.model01, subplot = 7)
```
Salida de las biomasas con todas las flotas

```{r}
SSplotTimeseries(base.model2, subplot = 7)
```

@@

## Residual

```{r eval=FALSE}
SSplotRunstest(base.model01, 
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model2, 
               subplots = "len",
               add=T)
```

```{r}
SSplotJABBAres(base.model2,
               subplots = "cpue",
               add=T)
SSplotJABBAres(base.model2,
               subplots = "len",
               add=T)
```

@@

```{r eval=FALSE}
retro(dir=dir7, oldsubdir="", 
      newsubdir="Retrospective", 
      years= 0:-5,
      exe="ss_osx",
      extras = "-nox", 
      skipfinished = F)
```

```{r echo=FALSE}
#s01
retroModels01 <- SSgetoutput(dirvec=file.path(dir01,
                                            "Retrospective",
                                            paste("retro",0:-5,
                                                  sep="")))

retroSummary01 <- SSsummarize(retroModels01)
# s1
retroModels1 <- SSgetoutput(dirvec=file.path(dir1,
                                            "Retrospective",
                                            paste("retro",0:-5,
                                                  sep="")))

retroSummary1 <- SSsummarize(retroModels1)
#s2
retroModels2 <- SSgetoutput(dirvec=file.path(dir2,
                                            "Retrospective",
                                            paste("retro",0:-5,
                                                  sep="")))

retroSummary2 <- SSsummarize(retroModels2)
#s3
retroModels3 <- SSgetoutput(dirvec=file.path(dir3,
                                            "Retrospective",
                                            paste("retro",0:-5,
                                                  sep="")))

retroSummary3 <- SSsummarize(retroModels3)
#s4
retroModels4 <- SSgetoutput(dirvec=file.path(dir4,
                                            "Retrospective",
                                            paste("retro",0:-5,
                                                  sep="")))

retroSummary4 <- SSsummarize(retroModels4)
#s5
retroModels5 <- SSgetoutput(dirvec=file.path(dir5,
                                            "Retrospective",
                                            paste("retro",0:-5,
                                                  sep="")))

retroSummary5 <- SSsummarize(retroModels5)
#s6
retroModels6 <- SSgetoutput(dirvec=file.path(dir6,
                                            "Retrospective",
                                            paste("retro",0:-5,
                                                  sep="")))

retroSummary6 <- SSsummarize(retroModels6)
#s7
retroModels7 <- SSgetoutput(dirvec=file.path(dir7,
                                            "Retrospective",
                                            paste("retro",0:-5,
                                                  sep="")))

retroSummary7 <- SSsummarize(retroModels7)

```

```{r echo FALSE}
#save(retroSummary, retroModels, file="retro5.Rdata")

retro01 <- SSplotRetro(retroSummary01,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)
retro1 <- SSplotRetro(retroSummary1,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)
retro2 <- SSplotRetro(retroSummary2,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)
retro3 <- SSplotRetro(retroSummary3,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)
retro4 <- SSplotRetro(retroSummary4,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)
retro5 <- SSplotRetro(retroSummary5,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)

retro6 <- SSplotRetro(retroSummary6,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)

retro7 <- SSplotRetro(retroSummary7,
            add=T,
            forecast = F,
            legend = T,
            verbose=F)

```

```{r echo=FALSE}
ggarrange(retro01,
          retro1,
          retro2,
          retro3,
          retro4,
          retro5,
          retro6,
          retro7, ncol = 4, nrow = 4)
```

@@

```{r}
hci = SSplotHCxval(retroSummary2, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
```
## Kobe (status)

```{r}
mvln = SSdeltaMVLN(base.model01,
                   Fref = "MSY", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

mvln = SSdeltaMVLN(base.model2,
                   Fref = "MSY", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)
```

```{r echo=FALSE}
# model01
tablebias01 <- SShcbias(retroSummary01,quant="SSB",verbose=F)
tablebias02 <- SShcbias(retroSummary01,quant="F",verbose=F)

kbl(tablebias01, booktabs = T,format = "latex",
    caption = "Rho parameter in SSB model s01")  %>% 
    kable_styling(latex_options = "HOLD_position")

kbl(tablebias02, booktabs = T,format = "latex",
    caption = "Rho parameter in F model s01")  %>% 
    kable_styling(latex_options = "HOLD_position")

# model2
tablebias7 <- SShcbias(retroSummary7,quant="SSB",verbose=F)
tablebias07 <- SShcbias(retroSummary7,quant="F",verbose=F)

kbl(tablebias7, booktabs = T,format = "latex",
    caption = "Rho parameter in SSB model s2")  %>% 
    kable_styling(latex_options = "HOLD_position")

kbl(tablebias7, booktabs = T,format = "latex",
    caption = "Rho parameter in F model s2")  %>% 
    kable_styling(latex_options = "HOLD_position")
```

@@

```{r eval=FALSE, echo = FALSE}
# 1. Identificar el directorio donde se encuentra el modelo base ----
dirname.model.run <- here("s2")
# 2. Crear un nuevo directorio para el "Perfil_Verosimilitud"  
dirname.R0.profile <- here("Ejercicios","Perfil_Verosimilitud")
dir.create(path=dirname.R0.profile, showWarnings = TRUE, recursive = TRUE)
# 3. Crear un subdirectorio llamado "plots_Verosimilitud" ----
plotdir=paste0(dirname.R0.profile, "/plots_Verosimilitud")
dir.create(path=plotdir, showWarnings = TRUE, recursive = TRUE)
# 4. Crear un subdirectorio llamado "simple" ----
reference.dir <- paste0(dirname.R0.profile,'/simple') 
dir.create(path=reference.dir, showWarnings = TRUE, recursive = TRUE)
# 5. Copiar el resultado del modelo base completo en este directorio ----
file.copy(Sys.glob(paste(dirname.model.run, "*.*", sep="/"),
                   dirmark = FALSE),
                    reference.dir)
# 6. Leer la salida del modelo base ----
Base <- SS_output(dir=reference.dir,covar=T)
# 7. Copiar los archivos necesarios de "simple" al directorio "Perfil_Verosimilitud" ----
copy_SS_inputs(dir.old = reference.dir, 
               dir.new =  dirname.R0.profile,
               copy_exe = TRUE,
               verbose = FALSE)
# 8. Leer los archivos del modelo ----
inputs <- r4ss::SS_read(dir = dirname.R0.profile)
# 9. Editar el archivo control la fase de estimación recdev ----
inputs$ctl$recdev_phase <- 1
# 10. Editar el archivo starter para leer los valores de inicio ----
inputs$start$init_values_src <- 0
# 11. Vector de valores para el perfil ----
R0.vec <- seq(15,20,0.5)  
Nprof.R0 <- length(R0.vec)
# 12. Cambiar el nombre del archivo control en el archivo starter.ss ----
inputs$start$ctlfile <- "control_modified.ss" 
# 13. Incluir prior_like para parámetros no estimados ----
inputs$start$prior_like <- 1                                 
# 14. Escribir los modelos modificados ----
r4ss::SS_write(inputs, dir = dirname.R0.profile, overwrite = TRUE)
# 15. Ejecutar la función profile() ----
?SS_profile()
profile <- profile(dir=dirname.R0.profile, # directory
                      exe="ss_osx",
                      oldctlfile ="control.ss",
                      newctlfile="control_modified.ss",
                      string="SR_LN(R0)",
                      profilevec=R0.vec)
# 16. Leer los archivos de salida ----
# (con nombres como Report1.sso, Report2.sso, etc.)
prof.R0.models <- SSgetoutput(dirvec=dirname.R0.profile, 
                              keyvec=1:Nprof.R0, 
                              getcovar = FALSE) 
# 17. Resumir las salidas con la función SSsummarize()  ----
prof.R0.summary <- SSsummarize(prof.R0.models)
# 18. Identificar los componentes de Verosimilitud ----
mainlike_components         <- c('TOTAL',
                                 "Survey", 
                                 'Length_comp',
                                 "Age_comp",
                                 'Size_at_age',
                                 'Recruitment') 
mainlike_components_labels  <- c('Total likelihood',
                                 'Index likelihood',
                                 'Length likelihood',
                                 "Age likelihood",
                                 'Size_at_age likelihood',
                                 'Recruitment likelihood') 
# 19. Funciones para generar plots de perfil de verosimilitud ----
### SSplotProfile() ----
png(file.path(plotdir,"R0_profile_plot.png"),
    width=7,
    height=4.5,
    res=300,
    units='in')
par(mar=c(5,4,1,1))
SSplotProfile(prof.R0.summary,           # summary object
              profile.string = "R0",     # substring of profile parameter
              profile.label=expression(log(italic(R)[0])), 
              ymax=150,minfraction = 0.001,
              pheight=4.5, 
              print=FALSE, 
              plotdir=plotdir, 
              components = mainlike_components, 
              component.labels = mainlike_components_labels,
              add_cutoff = TRUE,
              cutoff_prob = 0.95)

Baseval <- round(Base$parameters$Value[grep("R0",Base$parameters$Label)],2)
Baselab <- paste(Baseval,sep="")
axis(1,at=Baseval,label=Baselab)
abline(v = Baseval, lty=2)
dev.off()

### SSplotComparisons() ----
# Comparación de series de tiempo 
labs <- paste("SR_Ln(R0) = ",R0.vec)
labs[which(round(R0.vec,2)==Baseval)] <- paste("SR_Ln(R0) = ",
                                               Baseval,"(Base model)")

SSplotComparisons(prof.R0.summary,
                  legendlabels=labs,
                  pheight=4.5,png=TRUE,
                  plotdir=plotdir,
                  legendloc='bottomleft')

### PinerPlot() ----
#### R0_profile_plot_Length_like ----
png(file.path(plotdir,"R0_profile_plot_Length_like.png"),
    width=7,
    height=4.5,
    res=300,
    units='in')
par(mar=c(5,4,1,1))
PinerPlot(prof.R0.summary, 
          profile.string = "R0", 
          component = "Length_like",
          main = "Changes in length-composition likelihoods by fleet",
          add_cutoff = TRUE,
          cutoff_prob = 0.95)
Baseval <- round(Base$parameters$Value[grep("SR_LN",
                                      Base$parameters$Label)],2)
Baselab <- paste(Baseval,sep="")
axis(1,at=Baseval,
     label=Baselab)
abline(v = Baseval, lty=2)
dev.off()
#### R0_profile_plot_Survey_like ----
png(file.path(plotdir,"R0_profile_plot_Survey_like.png"),
    width=7,
    height=4.5,
    res=300,
    units='in')
par(mar=c(5,4,1,1))
PinerPlot(prof.R0.summary, 
          profile.string = "R0", 
          component = "Surv_like",
          main = "Changes in Index likelihoods by fleet",
          add_cutoff = TRUE,
          cutoff_prob = 0.95, legendloc="topleft")
Baseval <- round(Base$parameters$Value[grep("SR_LN",
                                            Base$parameters$Label)],2)
Baselab <- paste(Baseval,sep="")
axis(1,at=Baseval,label=Baselab)
abline(v = Baseval, lty=2)
dev.off()
```

@@

```{r echo=FALSE}
#PLOT labels, name of each model run
legend.labels <- c( "s01",
                  "s1",
                 "s2",
                 "s3",
                 "s4",
                 "s5",
                 "s6",
                 "s7")

#read in all model runs
#note if cover=T you need a hessian; if covar=F you do not need a hessian
biglist <- SSgetoutput(keyvec = NULL,
                       dirvec = c(dir01,
                                  dir1,
                                  dir2,
                                  dir3,
                                  dir4,
                                  dir5,
                                  dir6,
                                  dir7),
                       getcovar = F)

#create summary of model runs from list above
summaryoutput <- SSsummarize(biglist)

SSplotComparisons(summaryoutput,
                  legendlabels = c("s01",
                                   "s1",
                                   "s2",
                 "s3",
                 "s4",
                 "s5",
                 "s6",
                 "s7"),
                 labels = c("Year", 
                            "Spawning biomass (t)",
                            "Relative spawning biomass", 
                            "Age-0 recruits (1,000s)",
                            "Recruitment deviations", 
                            "Index", "Log index", 
                            "1 - SPR", 
                            "Density",
                            "Management target", 
                            "Minimum stock size threshold",
                            "Spawning output",
                            "Harvest rate"))

comp <- SSplotComparisons(summaryoutput,
                  legendlabels = c("s01",
                                   "s1",
                                   "s2",
                 "s3",
                 "s4",
                 "s5",
                 "s6",
                 "s7"),
                 subplots = c(2, 16))
comp
```

```{r}
comtable <- SStableComparisons(summaryoutput,
                   likenames = c("TOTAL", 
                                 "Survey", 
                                 "Length_comp",
                                 "Age_comp", 
                                 "priors",
                                 "Size_at_age"), 
                   names = c("Recr_Virgin",
                             "R0", 
                             "steep",
                             "NatM",
                             "L_at_Amax", 
                             "VonBert_K", 
                             "SSB_Virg", 
                             "Bratio_2017",
                             "SPRratio_2016"),
                   digits = NULL,
                   modelnames = c("s01", "s1","s2",
                 "s3",
                 "s4",
                 "s5",
                 "s6",
                 "s7"),
                   csv = FALSE,
                   csvdir =~"/DOCAS/SA_Krill",
                   csvfile = "parameter_comparison_table.csv",
                 verbose = TRUE,
                   mcmc = FALSE)
kbl(comtable, booktabs = T,format = "latex",
    caption = "Comparacion likelihood y parámetros s01, s1, s2, s3, s4, s5, s6 y s7")  %>% 
    kable_styling(latex_options = "scale_down")

```

@@

@@

```{r path, fig.cap="Framework path to stock assessment model in krill in WAP"}
knitr::include_graphics('Figs/pathmod.png')
```

Plot data

```{r datas2, out.width='60%', fig.cap="Data series using in "s2" scenario"}
knitr::include_graphics('s2/plots/data_plot.png')
```

\end{landscape}

\newpage
# REFERENCES

>>>>

```{r setup1, echo=FALSE}
rm(list = ls())
knitr::opts_chunk$set(eval=TRUE,
                      echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = 'center',
                      fig.pos = "H",
                      dev = 'jpeg',
                      dpi = 300,
                      tidy.opts=list(width.cutoff=50), 
                      tidy=TRUE)
#XQuartz is a mess, put this in your onload to default to cairo instead
options(bitmapType = "cairo") 
# (https://github.com/tidyverse/ggplot2/issues/2655)
# Lo mapas se hacen mas rapido
```

```{r message=FALSE, eval=TRUE}
library(here)
library(kableExtra)
library(ggpubr)
library(tibble)
```

@@

The stock assessment model for *Euphausia superba* (hereafter krill) in the Antarctic Peninsula, operating within a spatial-temporal and ecosystem framework, highlights the importance of incorporating spatial heterogeneity. The spatial heterogeneity, such as variations in oceanographic conditions and habitat suitability across different regions, is likely to outperform a population dynamics of krill. By accounting for spatial heterogeneity, a integrated stock assessment model can better capture localized dynamics and interactions, leading to more accurate predictions of krill population dynamics. This nuanced approach allows for a more refined understanding of different areas features krill structure, fishing mortality and recruitment patterns across the Antarctic Peninsula. Consequently, by acknowledging and integrating spatial heterogeneity, the stock assessment model can provide more robust insights into the ecological dynamics of krill populations and improve CCAMLR fishery management and conservation strategies in this region.

@@

Antarctic krill (*Euphausia superba* krill thereafter) population is predominantly concentrated in the Western Antarctic Peninsula (WAP), serving as a vital food source for predators and being targeted by increasing fishing activities. On another hand, WAP is considered one of the most sensitive areas to Climate Change (CC) and has already experienced fast changes in various dimensions [@Turner2005; @Lima2013; @McBride2021; @Morley2020; @Atkinson2019a]. Given its ecological significance and economic value, effective management of krill stocks is paramount to ensure the sustainability of Antarctic ecosystems and support fisheries-dependent communities. As part of its mandate, CCAMLR routinely conducts stock assessments to assess the status of important species, such as Antarctic krill, and to provide information for decision-making regarding fisheries management and conservation measures [@Hill2016]. This regional fisheries management organization has intensified its efforts to promote sustainable management of Antarctic krill through an ecosystem-based approach [@Watters2013].

Various stock assessments research on krill have been conducted to provide management advice to CCAMLR using model-based methodologies. However, many of these models face challenges in aligning with management schemes or accurately representing the population dynamics of krill, due to its variability and heterogeneity in spatial distribution and ecosystem complexity. Traditionally, stock assessments have relied on single-source data, primarily from fisheries catch data or population monitoring surveys. However, these approaches may overlook critical aspects of krill dynamics, such as spatial heterogeneity in abundance and distribution, as well as the influence of environmental factors and predator-prey interactions. Recognizing these limitations, there is a growing consensus on the need for integrated stock assessment models that incorporate multiple sources of information and consider the complex interplay of ecological variables. Integrating the complexity of krill population dynamics presents several challenges. Firstly, data consistency over the years has been problematic, as many monitoring programs and studies have not adhered to a uniform standard throughout their temporal and spatial extension. Additionally, certain regions lack sufficient data or possess only limited information regarding krill dynamics. Furthermore, interpreting indirect indicators that may affect the krill population, such as environmental variables, predator abundance, and fishery data, adds another layer of complexity. Despite these gaps, the existing data on krill population dynamics is abundant, considered data rich, and this availability allows build a integrated model to get populations signals and estimate main variables such as spawning biomass, recruitment, and fishing mortality, among others.

In response to this need, we propose a conceptual model to conducted a stock assessment of Antarctic krill in the Antarctic Peninsula, leveraging an integrated modeling approach that incorporates the fishing and biological history of krill over the past 20 years, as well as ecosystem variables such as environmental conditions, primary productivity levels affecting krill recruitment, and predator abundance. Unlike previous assessments, this apporach integrate data from both fisheries and population monitoring efforts, providing a more holistic understanding of krill dynamics. Moreover, our approach will go beyond conventional methods by explicitly considering spatial heterogeneity in krill abundance and distribution across the Antarctic Peninsula.

Spatial heterogeneity is a fundamental aspect of krill ecology, driven by a myriad of factors including oceanographic conditions and predator foraging behavior. By incorporating spatially implicit information into our model, we aim to capture the complex spatial patterns of krill distribution and better understand the underlying drivers of variability in krill abundance. This spatially implicit approach will allow us to identify krill hotspots, areas of high productivity, and potential ecological corridors, providing valuable insights for conservation planning and management strategies [@Watters2013]. Consequently, this approach aligns with current management frameworks, accurately reflecting krill population dynamics at a spatial scale and providing management recommendations at more localized scales than the current global ones.

Beyond integrating available information for assessing krill population dynamics, we tested several natural scenarios to identify uncertainties associated with estimating key population variables. Using a set of eight scenarios, we aimed to diagnose, evaluate, assess performance, and compare the tested models. This methodology represents a modern approach to assessing population dynamics and provides a foundation for making recommendations for the sustainable management of krill populations.

@@

The study area for the assessment of krill populations in the Antarctic Peninsula is block 48.1 subarea and strata subdivision inside to account for spatial heterogeneity. The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill. Therefore, dividing the study area into strata allows for a more detailed and accurate assessment of krill populations by considering local variations in environmental factors.

@@

In addition to existing data sources, researchers may submit requests to the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) Secretariat for access to additional data relevant to krill population assessment. These requests may include data from scientific research programs, environmental monitoring initiatives, and other sources that can contribute to improving the accuracy and reliability of the assessment. all this component and road to construct models, diagnostics and check outputs is represented in Figure \@ref(fig:path).

@@

Where:

\(N_t\) is the abundance of krill at time \(t\),  \(N_{t-1}\) is the abundance of krill at the previous time step, \(r\) is the intrinsic growth rate of the population, \(M\) is the mortality rate and \(R\) is the recruitment of new individuals into the population.

@@

The stock assessment model was configured using Stock Synthesis (SS3 hereafter)[SS3](https://vlab.noaa.gov/web/stock-synthesis) [@Methot2013; @methot2020stock] with the most updated version (V.3.30.21). SS3 is a structured age and size stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"*. SS3 has a stock population sub-model that simulates growth, maturity, fecundity, recruitment, movement, and mortality processes, and observation sub-models and expected values for different types of data. The model is coded in `C++` with estimation parameters enabled by automatic differentiation (ADMB) [@Fournier2012; @Methot2013: @methot2020stock]. The analysis of results and outputs uses R tools and the graphical interface of the *r4ss*  and *ss3diags* libraries [@Taylor2019; @Winker2023].

@@

## Environmental factor in modelation

One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of krill population dynamics. SS3 addresses these approaches. In the context of krill, the estimated level of recruitment depends on the spawning biomass from the previous season, an environmental time series—specifically, the 2000-2022 Chlorophyll-a time series—and a log-bias adjustment.

\[
E(\text{Recruitment}) = f(\text{SpBio}) \times \exp(B \times \text{envdata}) \times \exp\left(-0.5 \times \pi_R^2\right)
\]

\(R\) represents the variability of deviations, adding to the variance caused by environmental factors. Consequently, as the environmental effect accounts for more of the total recruitment variability, the residual \(R\) should be reduced. However, the model does not automatically adjust for this.

## Predator components in modelation

@@

In Table 1 we have eight scenarios to test different option in modeling about main consideration in assessment of krill population.

| Scenario | Description                                      |
|:---------:|:-------------------------------------------------|
|    s01     | Fishery, Predator, Survey, Environmental data agreggate (Whole 48.1) |
|    s1     | Fishery data (Length, Index, Catch) by strata  and Predator and Env data      |
|    s2    | Fishery and Survey (AMLR) data Length, Index, Catch by strata. Predator and Env data |
|    s3     | "s2" Without S-R relation         |
|    s4     | "s2" Ricker S-R relation estimated          |
|    s5     | "s2" BH S-R relation weak (0.9 steepness)     |
|    s6     | "s2" BH S-R relation strong (0.6 steepness)     |
|    s7     | "s2" BH S-R relation mid-strong (0.65 steepness) estimated     |

: Scenarios to modelling dynamics in krill in spatial implicit context

@@

@@

@@

## Kobe (status)

@@

@@

@@

 Some authors have idicated spatial heterogeneity and identified different zones, such as areas with higher or lower concentrations of juveniles or regions more affected by the environment, which structure the population differently even inside WAP (REF*).

@@

```{r path, fig.cap="Framework path to stock assessment model in krill in WAP"}
knitr::include_graphics('Figs/pathmod.png')
```
\end{landscape}

```{r concem, out.width='80%', fig.cap="Conceptual model used to model dynamics population in Antarctic krill in WAP"}
knitr::include_graphics('Figs/conceptual.png')
```

```{r datas2, out.width='60%', fig.cap="Data series using in s2 scenario"}
knitr::include_graphics('s2/plots/data_plot.png')
```

\begin{landscape}

```{r comparision, fig.cap="Data series using in s2 scenario"}
par(mfrow=c(2,2))
knitr::include_graphics(c('compare8_Fvalue_uncertainty.png',
                          'compare2_spawnbio_uncertainty.png',
                          'compare4_Bratio_uncertainty.png',
                          'compare8_Fvalue_uncertainty.png'))
```
\end{landscape}

# REFERENCES

### 2024-05-28 index.Rmd
improve Intro

In a simple way, the core of Stock Synthesis is its population dynamics model, which represents the dynamics of krill populations over time. This model incorporates key biological parameters such as growth rates, mortality rates, recruitment, and spawning biomass. The model is typically formulated using mathematical equations that describe how these parameters interact to determine the abundance and distribution of krill in the study area.

@@

```{r message=FALSE, eval=TRUE}
# install.packages("devtools")
# devtools::install_github("r4ss/r4ss", ref="development")

# 
# install.packages("caTools")
# library("caTools")
# 
# install.packages("r4ss")
library(r4ss)
library(here)
#remotes::install_github("PIFSCstockassessments/ss3diags")
library(ss3diags)
library(kableExtra)
library(doParallel) # facilita la ejecución paralela en R
detectCores()
registerDoParallel(8)
library(ggpubr)
```

@@

@@

>>>>

In a simple way, the core of Stock Synthesis is its population dynamics model, which represents the dynamics of krill populations over time. This model incorporates key biological, environmental and predator data sources. The model is typically formulated using mathematical equations that describe how these parameters interact to determine the abundance and distribution of krill in the study area.

@@

```{r message=FALSE, eval=TRUE}
# install.packages("devtools")
# devtools::install_github("r4ss/r4ss", ref="development")
# install.packages("caTools")
# library("caTools")
# install.packages("r4ss")
library(r4ss)
library(here)
#remotes::install_github("PIFSCstockassessments/ss3diags")
library(ss3diags)
library(kableExtra)
library(doParallel) # facilita la ejecución paralela en R
detectCores()
registerDoParallel(8)
library(ggpubr)
library(tibble)
```

@@

all this framework try to follow recommendations of @Carvalho2021b

@@

```{r}
like01 <- base.model01$likelihoods_used
like01$model <- rep("s01", nrow(like01))
like01 <- rownames_to_column(like01, var = "Description")
like1 <- base.model1$likelihoods_used
like1$model <- rep("s1", nrow(like1))
like1 <- rownames_to_column(like1, var = "Description")
like2 <- base.model1$likelihoods_used
like2$model <- rep("s2", nrow(like2))
like2 <- rownames_to_column(like2, var = "Description")
like3 <- base.model3$likelihoods_used
like3$model <- rep("s3", nrow(like3))
like3 <- rownames_to_column(like3, var = "Description")
like4 <- base.model4$likelihoods_used
like4$model <- rep("s4", nrow(like4))
like4 <- rownames_to_column(like4, var = "Description")
like5 <- base.model5$likelihoods_used
like5$model <- rep("s5", nrow(like5))
like5 <- rownames_to_column(like5, var = "Description")
like6 <- base.model6$likelihoods_used
like6$model <- rep("s6", nrow(like6))
like6 <- rownames_to_column(like6, var = "Description")
like7 <- base.model7$likelihoods_used
like7$model <- rep("s7", nrow(like7))
like7 <- rownames_to_column(like7, var = "Description")
totalike <- rbind(like01,
                  like1,
                   like2,
                   like3,
                   like4,
                   like5,
                   like6,
                   like7)

```{r}
ggplot(totalike)+
  geom_bar( aes(y = reorder(Description, values), x =values),
            stat = "identity", position = "dodge") +
  theme_few() +
  facet_wrap(.~model, 
             ncol=4)+
  labs(y="",
       x="Likelihood")+
  xlim(0, 1000)+
  theme(axis.text.x = element_text(angle = 90, 
                                   hjust = 1)) 
```
